Let's talk about Thurstone & Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-06-09 paper2.qmd
review and editing

Additionally, we construct the matrix $R$ to map each row of the $IA$ matrix with a corresponding row from the $JK$ matrix. This matrix has $n$ rows and $6$ columns, where $n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}$. Here, the term ${n_{I} \cdot n_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of stimuli generated by each pair of individuals in the population. Thus, we define the matrix as in Equation ([-@eq-mat12]). 
$$
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
$$ {#eq-mat12}

It is easier to visualize the structure of the previously defined vectors and matrices by considering an example. Assuming $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the vectors and matrices described in Equations ([-@eq-mat11]) and ([-@eq-mat12]) take the form as in Equation ([-@eq-mat13]).

$$
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
$$ {#eq-mat13}

@@

We can illustrate the structure of these vectors more clearly with an example. Suppose researchers exclude the second student, the second text from each student, and the third judge from the setup shown in Equation ([-@eq-mat13]). Given $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the resulting vectors would have the following structure:

$$
S_{I} = \begin{bmatrix}
1 \\
0 \\
1 \\
1 \\
1
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
1 \\
1 \\
0
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
1 \\
0 \\
0 \\
0 \\
1 \\
0 \\
1 \\
0 \\
1 \\
0 
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
1 \\
0 \\
0 \\
0 
\end{bmatrix}
$$ {#eq-mat24}

@@

Finally, we define the *sample mechanism* $S$ in Equation ([-@eq-mat25]), which maps each element of $S_{IA}$ to every element of $S_{JK}$. Each element $s_{(i,a,h,b,j,k)}$ is a binary value indicating the presence or absence of data rows in the matrix $R$ resulting from the sample mechanism, as in Equation ([-@eq-mat26]). Thus, the vector contains $n$ elements, matching the number of rows in $R$, and the sum of its elements represents the total data sample: $n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}$. Here, the term ${n^{s}_{I} \cdot n^{s}_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of sampled stimuli generated by each pair of sampled individuals.

$$
s_{(i,a,h,b,j,k)} = \begin{cases} 
1 & \text{if data elements } i,a,h,b,j,k \text{ from } R \text{ are sampled} \\
0 & \text{if data elements } h,i,a,b,j,k \text{ from } R \text{ are missing}
\end{cases}
$$ {#eq-mat26}

@@

<!-- 3. segway for next section -->
However, due to concerns about the practical feasibility of the comparison task [@Boonen_et_al_2020; ], CJ assessments rarely implement an exhaustive pairings of sampled judges, stimuli, and individuals. Thus, a realistic scenario must account for the fact that judges typically compare only a subset of stimuli authored by a sample of individuals.

@@

>>>>

Additionally, we construct the matrix $R$ to map each row of the $IA$ matrix with a corresponding row from the $JK$ matrix. This matrix has $n$ rows and $6$ columns, where $n = {n_{I} \cdot n_{A} \choose 2} \cdot n_{J} \cdot n_{K}$. Here, the term ${n_{I} \cdot n_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of stimuli generated by each pair of individuals in the population. Thus, we define the matrix as in Equation ([-@eq-mat12]).

$$
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 1 & 2 & 1 & n_{K} \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
i & a & h & b & j & k \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & 1 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
n_{I} & n_{A}-1 & n_{I} & n_{A} & n_{J} & n_{K}
\end{bmatrix}
$$ {#eq-mat12}

It is easier to visualize the structure of the previously defined vectors and matrices by considering an example. Assuming $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the vectors and matrices described in Equations ([-@eq-mat11]) and ([-@eq-mat12]) take the form as in Equation ([-@eq-mat13]).
$$
I = \begin{bmatrix}
1 \\
2 \\
3 \\
4 \\
5 
\end{bmatrix} ; \;
J = \begin{bmatrix}
1 \\
2 \\
3 
\end{bmatrix} ; \;
IA = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
2 & 1 \\
2 & 2 \\
3 & 1 \\
3 & 2 \\
4 & 1 \\
4 & 2 \\
5 & 1 \\
5 & 2 
\end{bmatrix} ; \;
JK = \begin{bmatrix}
1 & 1 \\
1 & 2 \\
1 & 3 \\
2 & 1 \\
2 & 2 \\
2 & 3 \\
3 & 1 \\
3 & 2 \\
3 & 3 
\end{bmatrix} ; \;
R = \begin{bmatrix}
1 & 1 & 1 & 2 & 1 & 1 \\
1 & 1 & 1 & 2 & 1 & 2 \\
1 & 1 & 1 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
1 & 1 & 5 & 2 & 1 & 1 \\
1 & 1 & 5 & 2 & 1 & 2 \\
1 & 1 & 5 & 2 & 1 & 3 \\
\vdots & \vdots & \vdots & \vdots & \vdots & \vdots\\
4 & 2 & 5 & 2 & 3 & 1 \\
4 & 2 & 5 & 2 & 3 & 2 \\
4 & 2 & 5 & 2 & 3 & 3 \\
5 & 1 & 5 & 2 & 3 & 1 \\
5 & 1 & 5 & 2 & 3 & 2 \\
5 & 1 & 5 & 2 & 3 & 3 
\end{bmatrix}
$$ {#eq-mat13}

@@

We can illustrate the structure of these vectors more clearly with an example. Suppose researchers exclude the second student, the second text from each student, and the third judge from the setup shown in Equation ([-@eq-mat13]). Given $n_{I} = 5$, $n_{A} = 2$, $n_{J} = 3$, and $n_{K} = 3$, the resulting vectors would have the following structure:
$$
S_{I} = \begin{bmatrix}
1 \\
0 \\
1 \\
1 \\
1
\end{bmatrix} ; \;
S_{J} = \begin{bmatrix}
1 \\
1 \\
0
\end{bmatrix} ; \;
S_{IA} = \begin{bmatrix}
1 \\
0 \\
0 \\
0 \\
1 \\
0 \\
1 \\
0 \\
1 \\
0 
\end{bmatrix} ; \;
S_{JK} = \begin{bmatrix}
1 \\
1 \\
1 \\
1 \\
1 \\
1 \\
0 \\
0 \\
0 
\end{bmatrix}
$$ {#eq-mat24}

@@

Finally, we define the *sample mechanism* $S$ in Equation ([-@eq-mat25]), which maps each element of $S_{IA}$ to every element of $S_{JK}$. Each element $s_{(i,a,h,b,j,k)}$ is a binary value indicating the presence or absence of data rows in the matrix $R$ resulting from the sample mechanism, as in Equation ([-@eq-mat26]). Thus, the vector contains $n$ elements, matching the number of rows in $R$, and the sum of its elements represents the total data sample: $n^{s} = \binom{n^{s}_{I} \cdot n^{s}_{A}}{2} \cdot n^{s}_{J} \cdot n^{s}_{K}$. Here, the term ${n^{s}_{I} \cdot n^{s}_{A} \choose 2}$ represents the binomial coefficient, which quantifies the total number of unique comparisons possible between every pair of sampled stimuli generated by each pair of sampled individuals.
$$
s_{(i,a,h,b,j,k)} = \begin{cases} 
1 & \text{if data elements } i,a,h,b,j,k \text{ from } R \text{ are sampled} \\
0 & \text{if data elements } h,i,a,b,j,k \text{ from } R \text{ are missing}
\end{cases}
$$ {#eq-mat26}

@@

@@

<!-- 3. segway for next section -->
However, due to concerns about the practical feasibility of the comparison task [@Boonen_et_al_2020], CJ assessments rarely implement an exhaustive pairings of sampled judges, stimuli, and individuals. Thus, a realistic scenario must account for the fact that judges typically compare only a subset of stimuli authored by a sample of individuals.
